### 2024-04-02 paper/paper.qmd

---
title: "Are Greener Spaces, Safer Places?"
subtitle: "Investigating the Link Between Green Spaces and Crime in Toronto"
author: 
  - Jena Shah
thanks: "Code and data are available at: https://github.com/shahjena/greenspace_safeplace.git."
date: today
date-format: long
abstract: "Abstract TBD."
format: pdf
number-sections: true
bibliography: references.bib
---

# Introduction

TBD

# Data {#sec-data}

TBD

# Model

TBD

## Model set-up

TBD

## Model justification

TBD

# Results

TBD

# Discussion

## First discussion point {#sec-first-point}

TBD

## Second discussion point

TBD

## Third discussion point

TBD

## Weaknesses and next steps

TBD

\newpage

\appendix

# Appendix {-}

# Datasheet

TBD

# Model details {#sec-model-details}

# Posterior predictive check

TBD

# Diagnostics

TBD

\newpage

# References

@citeR